---
title: "Case studyEDA"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2024-01-27"
---
```{r}
# Set up
library(dplyr)

data <- read.csv("~/Desktop/2024 Case Study Challenge/2024-srcsc-superlife-inforce-dataset.csv")
View(data)
```

```{r}
smoking_related <- subset(data, Smoker.Status == "S")

# Count the number of smoking-related policies
number_of_policies <- nrow(smoking_related)

number_of_policies
```
Number of policies that policyholder smoked is 61740.

```{r}

# Filter data for Smoker.Status = 'S'
smokers_data <- subset(data, Smoker.Status == "S")

# Calculate the count of each Cause of Death
cause_counts <- table(smokers_data$Cause.of.Death)

# Convert counts to proportions
cause_percentages <- prop.table(cause_counts) * 100

# Sort in descending order
sorted_causes <- sort(cause_percentages, decreasing = TRUE)

sorted_causes
```
I00-I99 is Diseases of the circulatory system. it occupies 55% of the cause of death among the policyholder who smoked.
I00-I99 include:    
I00-I02 Acute rheumatic fever   
I05-I09 Chronic rheumatic heart diseases    
I10-I15 Hypertensive diseases   
I20-I25 Ischaemic heart diseases    
I26-I28 Pulmonary heart disease and diseases of pulmonary circulation   
I30-I52 Other forms of heart disease    
I60-I69 Cerebrovascular diseases    
I70-I79 Diseases of arteries, arterioles and capillaries    
I80-I89 Diseases of veins, lymphatic vessels and lymph nodes, not elsewhere classified    
I95-I99 Other and unspecified disorders of the circulatory system   

Thus, from the interventions table, if we want to focus on smokers, we can choose:    
Heart Health Screenings          
Chronic Disease Management (these two are highly related to smokers)      
Incentives for Preventive Screenings  
Holistic Health Assessments        

```{r}
# Filter for Smoker.Status = 'S'
smokers_data <- subset(data, Smoker.Status == "S")

# Create a contingency table
underwriting_result <- table(smokers_data$Underwriting.Class)
urban_rural_result <- table(smokers_data$Urban.vs.Rural)

underwriting_result
urban_rural_result

```
```{r}
# Filter for rows where a death has occurred
deaths_data <- 
  data%>% 
  filter(Death.indicator == 1)

# Group by region and cause of death, then count occurrences
death_causes_by_region <- 
  deaths_data %>%
  group_by(Region, Cause.of.Death) %>%
  summarise(Deaths = n(), .groups = 'drop')


death_causes_by_region

```

